\Huge Conferences in the Era of Expensive Carbon

### 2019-06-05 main.tex
A couple of comments.

Indeed, scientific societies arguably have an especially important role to
play, since one of their main activities is organizing international
conferences, which produce significant greenhouse gas emissions,
particularly from air travel.   A single round-trip flight from
Philadelphia to Paris emits the equivalent of about 1.7 tons of carbon
dioxide (\COtwoE) per passenger~\cite{carboncalculator}\iflater\bcp{Trim
  citation?}\fi---a substantial fraction of the {\em total} yearly emissions
for an average resident of the US (16.5 tons) or Europe (7
tons)~\cite{emissions}.  \iflater\bcp{For comparison, ...  (put in some
  figures about the footprints of a few other things that people might think
  of reducing---driving, eating meat, upgrading their fridge...)}\fi{}
Moreover, these emissions have no near-term 
technological fix, since jet fuel is difficult to replace with
renewable energy sources~\cite{elec-air}. 
% \iflater\bcp{Trimmed this sentence: Reducing the \COtwoE{} footprint
% of conferences will require a reduction in the footprint of the attendees'
% air travel.}\fi

@@

Expecting that this era is coming, what should we rationally do now to
ensure that ACM, and computer-science research broadly, continue to thrive
in a world where air travel is more expensive?  Believing that it needs to
come as quickly as possible to avert disaster, what can we do to accelerate
the process?

@@

Beyond imposing a price on our carbon footprint (or waiting for one
to be imposed from outside), a sane response to climate change should
clearly also include taking steps to reduce the size of this footprint.  For
conferences, there are many options that one might
contemplate~\cite{ClimateCommitteReport}, from easy 
but incremental changes like making program committee meetings virtual
instead of physical and enabling virtual participation via live-streaming,
all the way to complete re-imagining of conferences and their role in
science (holding conferences completely in virtual spaces, replacing
conferences with journals, etc.), with some intriguing hybrid forms in the
middle (regional conferences, mega-conferences, conferences split into
two locations on different continents, etc.).  
%
However, before we undertake large changes, we need to understand the
situation better.  For most of these ideas, it is hard to confidently
predict either the potential benefits (in terms of reduced emissions) or the
potential costs (to scientific exchange, career development of young
scientists, etc.).  Indeed, a preliminary analysis of registration data from
a number of SIGPLAN conferences paints a rather complicated picture.
%
\bcp{Put the discussion of Figure 1 (and an accompanying figure
showing distance traveled to several instances of the same conference) here;
this illustrates the difficulty and the need for careful data gathering and
analysis.} 
All parts of ACM should be doing this kind of analysis.  (The acm-climate
mailing list is one place to exchange ideas about this.)

>>>>

Indeed, scientific societies arguably have an especially important role to
play, since one of their main activities is organizing international
conferences, which produce significant greenhouse gas emissions,
particularly from air travel. \mwh{The previous is the wrong
  transition, since again it seems to suggest that travel is a big
  deal in terms of the overall problem. The sentence also fails to
  convey the other reason scientific societies should be involved: As
  role models and advocates for general change.}  A single round-trip flight from
Philadelphia to Paris emits the equivalent of about 1.7 tons of carbon
dioxide (\COtwoE) per passenger~\cite{carboncalculator}\iflater\bcp{Trim
  citation?}\fi---a substantial fraction of the {\em total} yearly emissions
for an average resident of the US (16.5 tons) or Europe (7
tons)~\cite{emissions}.  \iflater\bcp{For comparison, ...  (put in some
  figures about the footprints of a few other things that people might think
  of reducing---driving, eating meat, upgrading their fridge...)}\fi{}
Moreover, these emissions have no near-term 
technological fix, since jet fuel is difficult to replace with
renewable energy sources~\cite{elec-air}. 
% \iflater\bcp{Trimmed this sentence: Reducing the \COtwoE{} footprint
% of conferences will require a reduction in the footprint of the attendees'
% air travel.}\fi

@@

Expecting that this era is coming, what should we rationally do now to
ensure that ACM, and computer-science research broadly, continue to thrive
in a world where air travel is more expensive?  Believing that it needs to
come as quickly as possible to avert disaster, what can we do to accelerate
the process? \mwh{The question now seems to have changed. The paper started by
  saying ACM has an ``especially important role to play,'' but now
  seems to suggest it should be reacting to an inevitable/coming
  future. These questions probably need to go much earlier, and the earlier
  text reworked around them.}

@@

Beyond imposing a price on our carbon footprint (or waiting for one
to be imposed from outside), a sane response to climate change should
clearly also include taking steps to reduce the size of this
footprint.
\mwh{Don't like previous sentence. It seems to have gotten away from
  the point of the paper, which is using price to change balance of
  apples to oranges. Analyzing data is useful for figuring out ACM's
  internal price, not tweaking around the edges, which is how it reads
  below.} For
conferences, there are many options that one might
contemplate~\cite{ClimateCommitteReport}, from easy 
but incremental changes like making program committee meetings virtual
instead of physical and enabling virtual participation via live-streaming,
all the way to complete re-imagining of conferences and their role in
science (holding conferences completely in virtual spaces, replacing
conferences with journals, etc.), with some intriguing hybrid forms in the
middle (regional conferences, mega-conferences, conferences split into
two locations on different continents, etc.).  
%
However, before we undertake large changes, we need to understand the
situation better.  For most of these ideas, it is hard to confidently
predict either the potential benefits (in terms of reduced emissions) or the
potential costs (to scientific exchange, career development of young
scientists, etc.).  Indeed, a preliminary analysis of registration data from
a number of SIGPLAN conferences paints a rather complicated picture.
%
\bcp{Put the discussion of Figure 1 (and an accompanying figure
showing distance traveled to several instances of the same conference) here;
this illustrates the difficulty and the need for careful data gathering and
analysis.} 
All parts of ACM should be doing this kind of analysis.  (The acm-climate
mailing list is one place to exchange ideas about this.)
